---
title: "Wed-project"
author: "C. A. Hamm"
date: "June 22, 2016"
output: html_document
---

## Analysis of vegetation diversity for the NEON D17 region
### Ultimately want to compare diversity estimates (alpha, beta, and gamma) between ground and remote data

```{r first-things}
set.seed(786121246)
setwd("~/Desktop/Projects/NEON-2016/Wed-Project/")

library("raster")
library("rgdal")
library("vegan")
library("vegetarian")
library("dplyr")
library("reshape")
options(stringAsFactors = FALSE)

# save and load data for future ease
# save(file = "SJER_SOAP_veg.RData", list = ls())
# load(file = "SJER_SOAP_veg.RData")
```

### Now to import the ground truthed data from D17

```{r import-data}
D17_veg <- read.csv("/Volumes/AOP-NEON1-4/D17/Field_Data/2013/Sampling_Data/D17_Vegetation_Structure/D17_2013_vegStr.csv", header = TRUE)
dim(D17_veg)
str(D17_veg)
head(D17_veg)
```

### Now to seperate data by site and plot ID, then generate an ecological count matrix

```{r parse-data}
# test run with only SJER data
CA_data <- D17_veg %>% group_by(plotid) %>% select(siteid, plotid, taxonid) 
head(CA_data)

# This will sum the number of each species idnetified in each plot 
CA_sites <- table(CA_data$plotid, CA_data$taxonid)
dim(CA_sites)

# Parse into SJER site
SJER_site <- CA_sites[1:18, ]
dim(SJER_site)
head(SJER_site)
tail(SJER_site)

# Parse into SOAP site
SOAP_site <- CA_sites[19:36, ]
dim(SOAP_site)
head(SOAP_site)
tail(SOAP_site)
```

### With our data parsed we can estimate D for the SJER and SOAP sites
# this is a one off to run the function automatically for and only use three arguements
```{r calculate-metrics}
D.one <- function(data, level, q){
		iter_q <- d(data, lev = level, q = q, boot = TRUE, boot.arg = list(num.iter = 1e3))
	return(iter_q)
}
D.one(data = SJER_site, level = "alpha", q = 0)


# Now a more general function to calculate for a range of values
D.iter.q <- function(data, level, q){
Spoon <- matrix(data = NA, ncol = 2, nrow = 6)
  for (i in 1:6){
  temp <- d(data, lev = level, q = i - 1, boot = TRUE, boot.arg = list(num.iter = 1e3))
  Spoon[i, 1] <- temp[[1]]
  Spoon[i, 2] <- temp[[2]]
  }
  return(Spoon)
}

# calculate the a, b, and g metrics for SJER
SJER.out.a <- D.iter.q(data = SJER_site, level = "alpha", q = 5)
SJER.out.b <- D.iter.q(data = SJER_site, level = "beta", q = 5)
SJER.out.g <- D.iter.q(data = SJER_site, level = "gamma", q = 5)

# calculate the a, b, and g metrics for SOAP
SOAP.out.a <- D.iter.q(data = SOAP_site, level = "alpha", q = 5)
SOAP.out.b <- D.iter.q(data = SOAP_site, level = "beta", q = 5)
SOAP.out.g <- D.iter.q(data = SOAP_site, level = "gamma", q = 5)
```

### Now plot the data and standard errors

```{r plot-metrics-alpha}
# a little function to plot
plotDeMoney <- function(x0, y0, x1, y1, mean, dot){
	points(y = mean, x = dot, pch = 15, col = 'grey', cex = 1.5)
  segments(x0, y0, x1, y1, lwd = 3.5)
}

# now create the plot for alpha at SJER
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 3, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(alpha))), las = 1, xlab = expression(paste(italic("q"))), main = "SJER")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SJER.out.a[i , 1] - SJER.out.a[i, 2], x1 = i - 1, y1 = SJER.out.a[i , 1] + SJER.out.a[i, 2], mean = SJER.out.a[i, 1], dot = i - 1)
}
    


# create plot for alpha diversity at SOAP
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 6, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(alpha))), las = 1, xlab = expression(paste(italic("q"))), main = "SOAP")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SOAP.out.a[i , 1] - SOAP.out.a[i, 2], x1 = i - 1, y1 = SOAP.out.a[i , 1] + SOAP.out.a[i, 2], mean = SOAP.out.a[i, 1], dot = i - 1)
}
```

### Create plots for beta diversity

```{r plot-metrics-beta}
# now create the plot for beta at SJER
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 4.2, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(beta))), las = 1, xlab = expression(paste(italic("q"))), main = "SJER")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SJER.out.b[i , 1] - SJER.out.b[i, 2], x1 = i - 1, y1 = SJER.out.b[i , 1] + SJER.out.b[i, 2], mean = SJER.out.b[i, 1], dot = i - 1)
}

# create plot for beta diversity at SOAP
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 6, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(beta))), las = 1, xlab = expression(paste(italic("q"))), main = "SOAP")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SOAP.out.b[i , 1] - SOAP.out.b[i, 2], x1 = i - 1, y1 = SOAP.out.b[i , 1] + SOAP.out.b[i, 2], mean = SOAP.out.b[i, 1], dot = i - 1)
}

```

### lastly the plots for gamma diversity

```{r plot-metrics-gamma}
# now create the plot for gamma at SJER
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 12, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(gamma))), las = 1, xlab = expression(paste(italic("q"))), main = "SJER")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SJER.out.g[i , 1] - SJER.out.g[i, 2], x1 = i - 1, y1 = SJER.out.g[i , 1] + SJER.out.g[i, 2], mean = SJER.out.g[i, 1], dot = i - 1)
}

# create the plot for gamma diversity at SOAP
plot(x = seq(from = 0, to = 5, length.out = 4), y = seq(from = 0, to = 18, length.out = 4), xaxt = "n", type = "n", ylab = expression(paste(italic(gamma))), las = 1, xlab = expression(paste(italic("q"))), main = "SJER")
axis(1, at = c(0, 1, 2, 3, 4, 5))
# populate the plot
for(i in 1:6){
  plotDeMoney(x0 = i - 1, y0 = SOAP.out.g[i , 1] - SOAP.out.g[i, 2], x1 = i - 1, y1 = SOAP.out.g[i , 1] + SOAP.out.g[i, 2], mean = SOAP.out.g[i, 1], dot = i - 1)
}
```